---
title: "BasicDiff"
author: "NithaDuff"
date: "7/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_dependencies}
library(dplyr)
library(ggplot2)
```
## Central Limit Theorem

n this project you will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations. You will investigate the distribution of averages of 40 exponentials. Note that you will need to do a thousand simulations.

```{r initialize}
lambda <- 0.2
mean <- sd <-  1/lambda
n <- 40
nosim <- 1000
x <- rexp(n,lambda)
```

```{r create_simulations}
samples <- matrix(sample(x,n*nosim,replace = T),nosim,n)
means <- apply(samples,1,mean)
vars <- apply(samples, 1, var)
```
The variation between the theoretical mean and the mean obtained from the simulation show a small variation which can be narrowed further on increasing either the sample size or the number or simulations.  
```{r comparison_mean}
mu0 <- mean(means)
print(c(mean,mu0))
```
The variance change from the theoretical value and the simulated value are also converging to the expected value.  
```{r comparison_var}
print(c(sd^2,mean(vars)))
```
##  Distribution of means

```{r distribution_mean}
hist(means, probability = TRUE)
curve(dnorm(x, mean=mean(means), sd=sd(means)), col="darkblue", lwd=2, add=TRUE, yaxt="n")
abline(v = mean, col = "red", lwd = 2)
abline(v = mean(means),col = "green", lwd = 2)
```

The above plot is overlayed with a curve that is approximately normal or is almost normal. The means of the two distributions vary by a certain quantity .The green line here shows the theoretical mean whereas the red line is the observed mean.

##  Distribution of variance  

```{r distribution_sd}

hist(vars, probability = TRUE)
curve(dnorm(x, mean=mean(vars), sd=sd(vars)), col="darkblue", lwd=2, add=TRUE, yaxt="n")
abline(v = sd^2, col = "red", lwd = 2)
abline(v = mean(vars),col = "green", lwd = 2)
```

The green line here shows the theoretical variance and the red line shows the observed variance.  
Moreover the distributions obtained resembles a normal distribution as shown by the overlayed curve.